Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-04 paper/lessons.tex
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7
More edits.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@605 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Lessons Learned}

@@

MapReduce algorithms can be hard to design, but are often relatively
easy to write and debug.  Thinking about a computational task from a
MapReduce perspective is different than traditional distributed-memory
parallel algorithm design.  For example, with an MPI mindset, it often
seems heretical to intentionally ignore data locality.  However,
writing small map() and reduce() functions is typically easy.  And
writing an algorithm that involves complex parallel operations,
without actually needing to write application-specific parallel code
to move data via MPI calls, is often a pleasant surprise.  Moreover,
if the MapReduce algorithm is initially coded so that it runs
correctly on one processor, it often works out-of-the-box on hundreds
or thousands of processors, without the need for additional debugging.

Performing MapReduce operations on a fixed allocation of processors on
a traditional MPI-based parallel machine is a somewhat different
conceptual model than that of cloud-computing MapReduces using (for
example) Hadoop.  In the former case, one can potentially control
which processor owns which data at various stages of an algorithm.
This is somewhat hidden from the user in a typical cloud-computing
model, where data simply exists somehere in the cloud, and Hadoop
ensures data moves where it needs to and is operated on by some
processor.  The cloud model is a nice data-centric abstraction which
allows for fault tolerance both to data loss (via redundant storage)
and to processor failure (via reassignment of work), neither of which
is typically possible on current MPI-based parallel machines.

@@

Finally, though this paper focuses on graph algorithms expressed as
MapReduce operations, there is nothing about the MR-MPI library itself
that is graph centric.  We hope the library can be generally useful on
large-scale monolithic or cloud-style parallel machines which support
MPI, for a variety of data-intense or compute-intense problems that
are amenable to solution using a MapReduce paradigm.

>>>>

\section{Lessons Learned}
\label{sec:lessons}

@@

MapReduce algorithms can be hard to design, but are often relatively
easy to write and debug.  Thinking about a computational task from a
MapReduce perspective is different from traditional distributed-memory
parallel algorithm design.  For example, with an MPI mindset, it often
seems heretical to intentionally ignore data locality.  However,
writing small {\it map()} and {\it reduce()} functions is typically easy.  And
writing an algorithm that involves complex parallel operations,
without actually needing to write application-specific parallel code
to move data via MPI calls, is often a pleasant surprise.  Moreover,
if the MapReduce algorithm is initially coded so that it runs
correctly on one processor, it often works out-of-the-box on hundreds
or thousands of processors, without the need for additional debugging.

Performing MapReduce operations on a fixed allocation of processors on
a traditional MPI-based parallel machine is a somewhat different
conceptual model than that of cloud-computing MapReduces using (for
example) Hadoop.  In the former case, one can potentially control
which processor owns which data at various stages of an algorithm.
This is somewhat hidden from the user in a typical cloud-computing
model, where data simply exists somewhere in the cloud and Hadoop
ensures data moves where it needs to and is operated on by some
processor.  The cloud model is a nice data-centric abstraction which
allows for fault tolerance both to data loss (via redundant storage)
and to processor failure (via reassignment of work), neither of which
is typically possible on current MPI-based parallel machines.

@@

Finally, though this paper focuses on graph algorithms expressed as
MapReduce operations, there is nothing about the MR-MPI library itself
that is graph centric.  We hope the library can be generally useful on
large-scale monolithic or cloud-style parallel machines that support
MPI, for a variety of data-intense or compute-intense problems that
are amenable to solution using a MapReduce paradigm.
